A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-19 MSRChallenge2016.tex
and, really, 4 pages

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC; and the least common
pattern is AOB. If we assume that these patterns never appear together, they
cover 14.78\% of buggy files. Kim et al. mentioned that the patterns that they
use cover almost 30\% of real patches~\cite{kim2013}. Although we only detect 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to expressively characterize human bug-fixing behavior.

@@

Some automatic program repair approaches mutate source code using predefined
mutation operators such as addition, deletion and replacement. In order to
generate patches that make more sense to human developers, it would be
beneficial to analyze how developers mutate source code to fix bugs.

\vspace{1ex}\textbf{Approach.}
Because direct diffs are difficult to identify, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another across
this dataset.  To do this, for each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

@@

\vspace{1ex}
\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, meaning that Do statements were
replaced by Assert statements 0.81\% of the times.

@@

There exist threats to the validity of our study. Since we are using a DSL
provided by Boa to implement our analysis, the correctness of our analysis
depends on our programs as well as Boa itself. To mitigate the risk of errors in
implementation, we open source our implementation on
Github~\footnote{\url{https://github.com/}}.\todo{FIX THIS NOW.}  Some threats
to our study arise from limitations of Boa. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem. Another big limitation is that Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions. Thus, our template
matching and analysis of code changes (by counting each statement kind) could be
inaccurate. However, our results should provide upper bounds.

We have characterized bug fixing changes in Java.
The findings of our study provide a set of useful insights for automatic
program repair tools in Java. It suggests that patterns proposed by the
state-of-the-art in generate and validate bug report for Java are not enough to
cover the entirety of bug fixes in Java. Adding more patterns is one solution,
but may not be a permanent one. Our finding on Java bug fixes characteristic
suggests that mutation based program repair may need to add new fields or
methods to fix bugs in Java and achieve a human-like written patch. A deeper
look suggests that it can also improve its search for a patch by having the knowledge of how
likely it is to replace a faulty statement by the different kinds of statements
available. Overall, our findings motivate additional study of the problem of
repair in Java, as assumptions that underlie approaches that target C are
unlikely to translate directly to this domain. 
% demonstrate that our knowledge of bug fixing
% properties in Java is still lacking and therefore more research on it are
% warranted.

>>>>

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC (changing a branch condition); and the least common
pattern is AOB (adding an array out of bound checker). If we conservatively assume that these patterns never appear together, they
cover 14.78\% of buggy files. In the dataset studied by Kim et al., the overall
10 patterns cover almost 30\% of real patches~\cite{kim2013}. Although we only analyze 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to expressively characterize human bug-fixing behavior.

@@

Some automatic program repair approaches seek generality by using
higher-granularity mutation operators such as statement-level addition, deletion and replacement. To
support the generation of high-quality patches, we 
next analyze how developers mutate source code to fix bugs at this granularity
level.

\subsection{Detection Approach}
Because direct diffs are difficult to identify on this dataset, we heuristically approximate the
extent to which one statement type appears to be ``replaced'' by another.
For each modified file, we count the number of
appearances of each statement type in the file pre- and post-commit.  We then
compare the results to see how many of each statement type was removed, and how
many inserted, to roughly characterize the types of replacement that happen at a
per-file level.
%   For each statement type (as tagged by the Boa infrastructure
% for Java), we process the results as follows: if number of occurrences of one
% statement type decreased post-fix and the number of another type increased, we
% say that the first statement type was replaced by the second statement type for
% that file. 
Note that this analysis doesn't distinguish the
replacement of the same statement kind, since we are counting the amount of
appearances of each statement kind.

@@

\subsection{Results}

\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, meaning that Do statements were
replaced by Assert statements 0.81\% of the times.

@@

\noindent\textbf{Threats to validity.} Since we are using a DSL
provided by Boa, the correctness of our analysis
depends on both our programs and Boa itself. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem.  To mitigate the risk of errors in
implementation, we open source our scripts~\footnote{\todo{FIX THIS NOW}}.
Because Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions, our template
matching and analysis of code changes (by counting each statement kind) only
provide estimates of behavior; we consider our results as informative
approximations.

The findings of our study provide useful insights for automatic program repair
tools in Java. It suggests that patterns proposed by the state-of-the-art in
generate and validate bug report for Java are insufficient cover the extent of
bug fixes in our dataset.  Adding more patterns is one solution, but may not be
a general one.  Our findings further suggest that
mutation based program repair may need to consider field or method insertion to
achieve human-comparability in patches.  A deeper look suggests that such
techniques may benefit from leveraging probabilistic knowledge of what types of
statements are commonly inserted, deleted, or replaced in Java bug-fixing commits.
Overall, our findings motivate additional study of 
repair in Java, as assumptions that underlie approaches that target C seem
unlikely to translate directly.

% demonstrate that our knowledge of bug fixing
% properties in Java is still lacking and therefore more research on it are
% warranted.
